---
title: "Progress Memo 2"
subtitle: |
  | Final Project 
  | Data Science 1 with R (STAT 301-1)
author: "Nina Gupta"
date: today
pagetitle: "PM2 Nina Gupta"

format:
  html:
    toc: true
    toc-depth: 4
    toc-location: left
    embed-resources: true
    link-external-newwindow: true

execute:
  warning: false
  echo: false

from: markdown+emoji
reference-location: margin
citation-location: margin
---

::: {.callout-tip icon=false}

## Github Repo Link

[Nina's Final Project Memo 2 Github Repo](https://github.com/ninag0104/final-project-2)

:::


```{r}
#| label: load-data
#| results: hide

library(tidyverse)
library(naniar)
chicago_crimes <- read_csv("data/chicago_crimes.csv") |>
    janitor::clean_names() 
```


## Data Source Summary

[Chicago Crime Data](https://www.chicago.gov/city/en/dataset/crime.html)

As a reminder, the data source I chose is the City of Chicago's^[data collected by
Chicago Police Department] "Chicago Crime Data", presenting data from one year
prior to the current date, minus the last 7 days. The link below provides more
information on the source, data reliability, and variable meanings. 

[Description of Chicago Crime Data](https://data.cityofchicago.org/Public-Safety/Crimes-One-year-prior-to-present/x2n5-8w5q/about_data)

## Cleaning Data
To clean the data, I used the clean_names() function to convert the variable names
to all lowercase and add underscores for better readability. 

Considering I am exploring the relationship between location and crime committed, 
I decided to remove some identification variables I am not considering/that will 
not impact the analysis needed to answer my research question, including
iucr^[Illinois Uniform Reporting Code] and fbi_cd^[FBI's classification of crime].

Since I've already explored the missingness in Progress Memo 1 and determined
that all missing data relates to location, I've decided to filter out the missing 
data so that it doesn't negatively impact my analysis. 

```{r}
#| label: missingness-tbl

missing_variables <- chicago_crimes |>
  summarize(
    missing_location = sum(is.na(location)), 
    missing_location_description = sum(is.na(location_description)),
    missing_longitude = sum(is.na(longitude)), 
    missing_latitude = sum(is.na(latitude)), 
    missing_x_coordinate = sum(is.na(x_coordinate)),
    missing_y_coordinate = sum(is.na(y_coordinate)),
    missing_total = sum(is.na(chicago_crimes))
  ) |>
    pivot_longer(
    cols = everything(), 
    names_to = "variable", 
    values_to = "missing_count"
  )

knitr::kable(missing_variables, caption = "Table 1. Missingness of Specific
             Variables in Chicago Crime Data 
             (further exploring data from
             Figure 2)")

```


```{r}
#| results: hide

chicago_crimes_clean <- chicago_crimes |>
  janitor::clean_names() |>
  select(-iucr, -fbi_cd) |>
  na.omit(crime_crimes)
```


## EDA
After cleaning up my data by standardizing it, removing unnecessary variables, and
removing missing values, I am ready to complete EDA. To start, my initial research
question is: **What is the correlation between types of crimes committed and
location in the city of Chicago?**

## Univariate & Bivariate Analysis - graphics/tables w description of what has
## been found thus far. What exploring & why Share research question per graphic

## Summarize Progress (where at, next steps, self-assessment, guiding curiosities
## or research questions planned to explore)





PROGRESS MEMO 1: 

This dataset includes 257,912 observations and 17 variables containing both
numerical (location (coordinates), date, case number, etc.) and categorical 
variables (primary description (type of crime), secondary description (weapons
used), block where the crime was committed, etc). There are 6 numerical variables
and 11 categorical. With the amount of data, there are many relationships to be
explored in this dataset. 

I've explored the data using visuals that are in the "References" section of this
document^[Located at the bottom of the document]

```{r}
#| label: exploring-variables
#| results: hide

chicago_crimes_summary <- chicago_crimes |>
  na.omit(crime_crimes)

chicago_crimes_summary |>
  mutate(
    missingness = sum(is.na(chicago_crimes_summary)))
```

## Why this data

I chose this data because I was curious to learn more about crime rates in Chicago
and explore trends within the dataset, also considering we are located close to 
the city. 

Although I am still narrowing it down, I am interested exploring the
correlation between types of crimes committed and another variable, possibly
location or date of occurrence of the crime in Chicago. Since I am sure I want to 
explore the type of crime vs another variable, in Figure 1 I explored the
distribution of the types of crimes committed in Chicago.

## Data quality & complexity check

This data was collected by the City of Chicago during a specific time period^[one
year prior to current date minus the past 7 days]. Data is extracted from the
Chicago Police Department's CLEAR (Citizen Law Enforcement Analysis and Reporting)
system and is a reputable source of information, other than room for bias of 
police officers who collected the data. This is something I will have to note
while making conclusions about the data.

This dataset is updated daily, Tuesday through Sunday. This dataset is complex
considering the number of variables and distribution of types of crime^[as seen
in Figure 1] and that there are 6 variables with missing data, as 
shown in Figure 2 and Table 1. 

## Potential data issues

One issue with the data mentioned earlier is that the police officers who
collected it are subject to bias, which could be reflected in the data by
representing potentially harmful stereotypes and generalizations. Considering
crime is a complex and sensitive topic to explore, many incorrect assumptions can
be made without proper consideration of the influence of bias.

Another possible issue is missingness in the data, since there are 1589 values
in total missing in the dataset as seen in Table 1. The missing data is important
if I decide to focus on the relationship between type of crimes committed and
location, as the variables with missing data are x-coordinate, y-coordinate,
latitude and longitude, location, and location description. The exact counts are
shown in Table 1 and the variables all explore the location where the crime
occurred. 

This indicates it may be ideal to explore a variable other than location, such
as date or number of arrests, and its relation to the type of crime. In Figure 1,
I explored the distribution of the types of crimes using the primary_description
variable. 

## References


```{r}
#| label: fig-description-distribution
#| fig-cap: "Distribution of Crime Type"

na.omit(chicago_crimes) |> 
  ggplot(aes(primary_description)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + 
  labs(
    title = "Distribution of Crime Type",
    subtitle = "Displayed without primary_description NA values",
    x = "Crime Type", 
    y = "Count"
  )
```

@fig-description-distribution



```{r}
#| label: fig-missingness
#| fig-cap: "General Missingness in Chicago Crimes Data"

chicago_crimes |>
  gg_miss_var() + 
  labs(
    title = "General Missingness in Chicago Crimes Data", 
    subtitle = "Collected by Chicago Police Department's CLEAR", 
  )
```

@fig-missingness
